LV2

### 2018-05-10 paper.tex
Topics
First complete draft

Over the period from 15 April 2017 to \hl{19 April 2018} the initial set of \hl{1\,685\,048} tweets was collected from the \texttt{POST status/filter} endpoint of the Twitter Streaming API.\footnotemark{}
%
\footnotetext{\url{https://developer.twitter.com/en/docs/tweets/filter-realtime/api-reference/post-statuses-filter}}
%
The \texttt{locations} parameter was set to the bounding box of Riga, the capital of Latvia.\footnotemark{}
%
\footnotetext{The coordinates of the bounding box are \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}.}
%
In addition to the location, \hl{420} accounts were tracked. \hl{The whole list of screen names together with user ID is available in the supplement.}

@@

To investigate the collection further nine topics of interests were defined. The collection was manually searched using keywords to get relevant tweets. The topics are
\begin{itemize}
\item \texttt{LV001: Winter} Everything about winter. The relevant tweets are mostly about the weather, the pictures of snow and driving conditions.
\item \texttt{LV002: Mel\=anijas hronika} Mel\=anijas hronika is a Latvian movie that was selected for Foreign-Language Category for Oscars.
\item \texttt{LV003: Refugees} A broad topic about refugee crisis in Europe, attitude to immigration and immigrants.
\item \texttt{LV004: Dv\=eese\c{l}u putenis} A Latvian movie that is currently in production.
\item \texttt{LV005: Hockey} Ice hockey related tweets.
\item \texttt{LV006: Christmas} Christmas.
\item \texttt{LV007: New Year} New Year.
\item \texttt{LV008: Olympics} Tweets about the Olympic games and the Latvian team.
\item \texttt{LV009: Brexit} Brexit.
\end{itemize}

>>>>

Over the period from 15 April 2017 to \hl{19 April 2018} the initial set of \hl{1\,685\,048} tweets was collected from the \texttt{POST status/filter} endpoint of the Twitter Streaming API.\footnotemark{}
%
\footnotetext{\url{https://developer.twitter.com/en/docs/tweets/filter-realtime/api-reference/post-statuses-filter}}
%
The \texttt{locations} parameter was set to the bounding box of Riga, the capital of Latvia.\footnotemark{}
%
\footnotetext{The coordinates of the bounding box are \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}.}
%
In addition to the location, \hl{420} accounts were tracked. %\hl{The whole list of screen names together with user ID is available in the supplement.}

@@

To investigate the collection further nine topics of interests were defined. The collection was manually searched using keywords to get relevant tweets. The topics are:
\begin{itemize}
\item \texttt{LV001: Winter} Everything about winter.
\item \texttt{LV002: Mel\=anijas hronika} Mel\=anijas hronika is a Latvian movie that was selected for Foreign-Language Category for Oscars.
\item \texttt{LV003: Refugees} A broad topic about refugee crisis in Europe, attitude to immigration and immigrants.
\item \texttt{LV004: Dv\=eese\c{l}u putenis} A Latvian movie that is currently in production.
\item \texttt{LV005: Ice hockey} Ice hockey related tweets.
\item \texttt{LV006: Christmas} Christmas.
\item \texttt{LV007: New Year} New Year.
\item \texttt{LV008: Olympics} Tweets about the Olympic games and the Latvian team.
\item \texttt{LV009: Brexit} Brexit.
\end{itemize}

Topics exhibit various temporal patterns, refer to Figure~\ref{fig:topic_timeline}. The volume of tweets is constant for long-lasting news stories such as the refugee crisis, Brexit and ice hockey.
%
During the Ice Hockey World Championship the volume of hockey related tweets increases. Similarly, event-based topics---such as Christmas, New Year and Olympics---are the most active during corresponding events, the volume of tweets slowly builds up as an event approaches.

In case of Christmas, we see unexpected activity in March which is due to the discussion of making the Orthodox Christmas a public holiday in Parliament. This topic exhibits some cultural differences and language use. Tweets in Latvian peak during Advent and Christmas in December, tweets in Russian reach maximum both in December and early January when Orthodox Christmas are celebrated. It worth noting that during the discussion whether Orthodox Christmas should be a public holiday, it sometimes was referred as ``Russian Christmas''.

Topics about movies are the smallest volume-wise. For a movie that was shown at several international festivals, we see multilingual content, while tweets about a movie that is still in production are solely monolingual.

The topic about Winter is an example of a seasonal topic, which is mostly about the weather, the pictures of snow and driving conditions.

Table~\ref{tab:topic_lang_counts} shows the total number of tweets per topic and language share for Latvian, Russian and English. For all topics the share of Latvian tweets is higher than on average in the collection (Section~\ref{sec:language}). For Russian and English these numbers are generally lower than global average.

The most notable exception being Russian tweets about New Year, probably because of the content written by Russian tourists. The topic about Winter comes close for Russian with respect to the expected average, but is low for English tweets. Interestingly, for the international news topics, there is proportionally more English content about Brexit than about refugees.

\input{figures/topic_lang_counts.tex}

\section{Conclusion and future work}
\label{sec:conclusion}

This paper presents a multilingual tweet collection with some analysis of users, language use and content. The analysis reveal that the resource might be used in several areas ranging from computational linguistics, information retrieval and social science.

As for computational linguistics, multilingual language identification needs to be applied to it, as tweets are not always monolingual, as we indirectly showed by the large volume of English tweets coming from Instagram and Latvian tweets coming from Foursquare.

An evidence of the existence of reliable topics that exhibit various temporal patterns could be exploited in information retrieval either for evaluation of multilingual retrieval systems or development of collection exploration techniques.

For social analysis, it would be nice to see what topics are multilingual, for example Winter, and what topics are monolingual, for example movie related topics in this work, and why the topics are discussed differently depending on a language.

\input{figures/topic_timeline.tex}
